---
title: "Bisymmetric Bar Modelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bisymmetric Bar Modelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The bisymmetric module is for a galaxy with independent evidence of a bar. It
extends the axisymmetric disc model with the two-sided streaming terms,
following the velocity-field formulation used in the barred-galaxy literature.
It is not a generic substitute for the disc model.

## Before fitting

Capivara first estimates a photometric bar-angle prior from the central
white-light elongation, then deprojects it relative to the kinematic disc
major axis. It deliberately does not substitute the disc position angle: doing
so silently assumes an aligned bar and gives a misleading model for most
galaxies. Inspect the magenta prior in the component figure before interpreting
the bisymmetric terms. A measured in-plane angle can always override the
automatic estimate with `bar_phi_deg`.

## Fit the bar module

```{r bisymmetric-model, eval = FALSE}
library(capivara)

bar_result <- run_kinematic_analysis(
  cube_path = "/data/known_barred_galaxy.fits",
  redshift = 0.03,
  emission_line = "halpha",
  segmentation_mode = "kinematic",
  model = "bisymmetric_bar",
  knn_k = 50,
  n_segments = 25,
  model_control = list(
    disc_inc_deg = 38,
    robust_fit = TRUE
  ),
  show_plots = TRUE
)

plot(bar_result, which = "components")
```

The `components` plot is intentionally unavailable for an axisymmetric run:
the circular, tangential second-order, and radial second-order components only
have a physical interpretation within the bisymmetric model.

## Optional bar-support prior

The model normally uses the full valid kinematic footprint. When a carefully
measured bar footprint is needed as a weak geometric prior, add:

```{r bar-support, eval = FALSE}
model_control = list(
  bar_phi_deg = 41, # optional measured in-plane override
  use_bar_support_mask = TRUE,
  bar_support_width_deg = 25
)
```

This option should be compared against the full-footprint model, not used to
force a bar-like residual pattern.
