---
title: "Bar Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bar Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The bisymmetric module extends the axisymmetric disc model with two-sided
streaming terms. Use it for a galaxy with independent evidence of a bar.

## Before fitting

Capivara estimates a bar-angle prior from the central white-light elongation
and deprojects it relative to the kinematic disc major axis. It does not use the
disc position angle as the bar angle, because that would assume alignment. Use
`bar_phi_deg` to supply a measured in-plane angle.

## Fit the bar module

```{r bisymmetric-model, eval = FALSE}
library(capivara)

bar_result <- run_kinematic_analysis(
  cube_path = "/data/known_barred_galaxy.fits",
  redshift = 0.03,
  emission_line = "halpha",
  segmentation_mode = "kinematic",
  model = "bisymmetric_bar",
  knn_k = 50,
  n_segments = 25,
  model_control = list(
    disc_inc_deg = 38,
    robust_fit = TRUE
  ),
  show_plots = TRUE
)

plot(bar_result, which = "components")
```

The `components` plot is available only for the bisymmetric model. It shows the
circular, tangential second-order, and radial second-order terms.

## Optional bar-support prior

The model normally uses the full kinematic footprint. To add a measured bar
footprint as a weak geometric prior:

```{r bar-support, eval = FALSE}
model_control = list(
  bar_phi_deg = 41, # optional measured in-plane override
  use_bar_support_mask = TRUE,
  bar_support_width_deg = 25
)
```

Compare this result with the full-footprint model.

Start with the axisymmetric model in [Kinematic
Analysis](kinematic-analysis.html). Interpret a bar only after comparing the
model diagnostics with independent observations.
