---
title: "Segmentation Backends"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Segmentation Backends}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
```

Calls without an explicit representation retain Capivara's historical exact
Ward and sparse-graph Ward backends. They return compatible region maps and
regional-product inputs, but they make different computational approximations.
An explicit `full_flux` or `spectral_shape` declaration selects the
four-neighbour observed-entry hierarchy described in the Semantic
Representations vignette; both public segmentation functions then invoke the
same representation-specific backend.

## Exact Ward

`segment()` constructs the full pairwise distance object among valid spaxels.
The distance storage therefore grows quadratically with the number of eligible
spaxels. Estimate the lower-bound memory before a large run:

```{r exact-memory, eval=FALSE}
estimate_segment_memory(cube)

seg <- segment(
  input = cube,
  Ncomp = 25,
  use_starlet_mask = TRUE
)
```

Use the exact backend when its all-pairs distance matrix fits in memory.

## Sparse Ward

`segment_large()` avoids storing the full all-pairs distance matrix by building
a coherent nearest-neighbour graph.

```{r sparse-backend, eval=FALSE}
seg <- segment_large(
  input = cube,
  Ncomp = 25,
  use_starlet_mask = TRUE,
  knn_k = 100,
  auto_k = FALSE,
  verbose = TRUE
)

seg$backend_info
```

`knn_k` controls graph density. With `auto_k = TRUE`, Capivara may increase it
to connect the graph; the value used is returned in `backend_info`.

## Select spectral channels

By default, all wavelength channels drive the segmentation. Set
`feature_wavelength_range` to learn labels from a chosen interval while keeping
the full cube in the returned object:

```{r feature-window, eval=FALSE}
seg_window <- segment_large(
  input = cube,
  Ncomp = 25,
  feature_wavelength_range = c(lambda_min, lambda_max),
  use_starlet_mask = TRUE,
  knn_k = 100
)

full_regional_spectra <- summarize_cluster_spectra(seg_window)$sum_spectra
```

`feature_wavelength_range` selects the clustering channels.
`wavelength_range` selects the channels used by the S/N screen. The two
parameters are not aliases.

## Choose the component count

`choose_ncomp_by_snr()` evaluates candidate cuts and returns the largest tested
count whose minimum regional S/N remains above the requested threshold.

```{r snr-choice, eval=FALSE}
choice <- choose_ncomp_by_snr(
  input = cube,
  target_snr = 30,
  var_cube = variance_cube,
  k_values = c(5, 10, 15, 20),
  wavelength_range = c(lambda_min, lambda_max)
)
```

The result depends on the target, variance cube, candidate grid, and wavelength
interval. Representation-specific hierarchies currently retain `Ncomp` only as
a chronological compatibility cut; evidence-driven stopping is outside V4.
