---
title: "Choosing a Segmentation Backend"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing a Segmentation Backend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
```

Capivara exposes an exact Ward backend and a sparse-graph Ward backend. They
return compatible region maps and regional-product inputs, but they make
different computational approximations.

## Exact Ward

`segment()` constructs the full pairwise distance object among valid spaxels.
The distance storage therefore grows quadratically with the number of eligible
spaxels. Estimate the lower-bound memory before a large run:

```{r exact-memory, eval=FALSE}
estimate_segment_memory(cube)

seg <- segment(
  input = cube,
  Ncomp = 25,
  use_starlet_mask = TRUE
)
```

Use the exact backend when the supported footprint is small enough for the
available memory and the all-pairs Ward construction is part of the intended
analysis.

## Sparse Ward

`segment_large()` avoids storing the full all-pairs distance matrix by building
a coherent nearest-neighbour graph.

```{r sparse-backend, eval=FALSE}
seg <- segment_large(
  input = cube,
  Ncomp = 25,
  use_starlet_mask = TRUE,
  knn_k = 100,
  auto_k = FALSE,
  verbose = TRUE
)

seg$backend_info
```

`knn_k` controls graph density. It is an analysis parameter, not a hidden
performance toggle: record it with the input cube, support definition, random
seed where applicable, and output path. `auto_k` may increase graph density to
obtain a connected graph; inspect and retain `backend_info` when using it.

## Select spectral channels

By default, all wavelength channels drive the segmentation. Set
`feature_wavelength_range` to learn labels from a chosen interval while keeping
the full cube in the returned object:

```{r feature-window, eval=FALSE}
seg_window <- segment_large(
  input = cube,
  Ncomp = 25,
  feature_wavelength_range = c(lambda_min, lambda_max),
  use_starlet_mask = TRUE,
  knn_k = 100
)

full_regional_spectra <- summarize_cluster_spectra(seg_window)$sum_spectra
```

`feature_wavelength_range` selects clustering features. In contrast, when
`target_snr` is used, `wavelength_range` selects the channels used by the S/N
screen. The two intervals answer different questions and should not be treated
as aliases.

## Choose the component count

`choose_ncomp_by_snr()` evaluates candidate cuts and returns the largest tested
count whose minimum regional S/N remains above the requested threshold.

```{r snr-choice, eval=FALSE}
choice <- choose_ncomp_by_snr(
  input = cube,
  target_snr = 30,
  var_cube = variance_cube,
  k_values = c(5, 10, 15, 20),
  wavelength_range = c(lambda_min, lambda_max)
)
```

The target, variance treatment, candidate grid, and wavelength interval are
part of the scientific configuration and should accompany the result.
