---
title: "Regional Spectra"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regional Spectra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
```

The segmentation map is categorical. Summing the spaxels in each region gives
the flux-preserving spectrum used for fitting.

## Regional summaries

```{r summaries, eval=FALSE}
products <- summarize_cluster_spectra(
  seg,
  var_cube = variance_cube,
  variance_inflation = 1
)

products$cluster_ids
products$n_spaxels
products$sum_spectra
products$sum_variance
```

For region (k), channel \(\lambda\), and assigned spaxels \(p\), Capivara
computes

\[
F_{k,\lambda}^{\mathrm{sum}} = \sum_{p \in k} F_{p,\lambda}.
\]

When a variance cube is supplied, the regional variance is the sum of valid
input variances multiplied by `variance_inflation`. Capivara does not estimate
spatial covariance from the flux cube.

Means and medians describe spectral shape; only the summed spectra preserve
the total regional flux.

## Representative cubes

```{r reconstructions, eval=FALSE}
representative <- reconstruct_cluster_cube(
  seg,
  template = "median",
  preserve_mask = TRUE,
  return_residual = TRUE
)

flux_preserving <- reconstruct_flux_preserving_cube(
  seg,
  fill_mode = "na",
  return_residual = TRUE
)
```

`reconstruct_cluster_cube()` fills a region with a representative spectrum.
`reconstruct_flux_preserving_cube()` distributes each summed regional spectrum
over its assigned spaxels so summing the reconstruction over the region
recovers the regional total.

## Map–table correspondence

Region identifier `k` in `seg$cluster_map` corresponds to `k` in
`products$cluster_ids` and to the matching row of the returned spectral
matrices. Region identifiers remain categorical even when stored as integers.
