---
title: "Kinematic Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Kinematic Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(capivara)
```

Capivara separates three questions that are often mixed together:

1. Which spaxels belong together in emission-line kinematics?
2. Is an axisymmetric rotating disc an adequate comparison model?
3. Is there evidence for a more specific perturbation, such as a bar?

The first two are useful for both barred and unbarred galaxies. A bar model is
therefore not a default setting.

## 1. Kinematic-aware segmentation

`segment_kinematics()` makes the native line maps and clusters flux, velocity,
dispersion, and profile-shape information. It does not run ordinary
full-spectrum segmentation, which answers a different question.

```{r kinematic-segmentation, eval = FALSE}
library(capivara)

segments <- segment_kinematics(
  cube_path = "/data/galaxy.fits",
  redshift = 0.03,
  emission_line = "halpha",
  segmentation_mode = "kinematic",
  knn_k = 50,
  n_segments = 25,
  show_plots = TRUE
)
```

Choose `segmentation_mode = "path_signature"` when the path-signature feature
representation is part of the science question. It additionally writes that
segmentation; it does not replace the native kinematic maps.

`support_mode = "starlet"` is the conservative default and retains the
white-light galaxy footprint. For emission-line work in a field with appreciable
sky, set `support_mode = "line_flux"`; Capivara then uses a robust
border-noise threshold on the measured line-flux map and keeps its main
connected component. `line_flux_sigma = 3` is a useful strict starting point;
lower it if faint extended line emission is scientifically important.

## 2. Axisymmetric disc comparison

The default model is an axisymmetric disc. It produces an observed velocity
map, disc model, residual map, and circular-speed profile without making any
claim that the galaxy hosts a bar.

```{r axisymmetric-model, eval = FALSE}
result <- run_kinematic_analysis(
  cube_path = "/data/galaxy.fits",
  redshift = 0.03,
  emission_line = "halpha",
  model = "axisymmetric",
  segmentation_mode = "kinematic",
  knn_k = 50,
  n_segments = 25,
  show_plots = TRUE
)

plot(result, which = "model")
```

For a MaNGA LOGCUBE, leave `redshift = NA_real_` to resolve it from the local
metadata/header. For another IFU cube, provide the redshift explicitly.

## 3. Model modules

Use `kinematic_models()` to see the installed dynamical modules. Each module
uses the same native maps and kinematic segmentation, so new models such as a
spiral perturbation can be added without changing the upstream workflow.

```{r available-models}
kinematic_models()
```

The bar-specific workflow is documented separately in `vignette("bisymmetric-bar-model", package = "capivara")`.
