---
title: "Semantic Representations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Semantic Representations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
```

A spectral partition inherits the meaning of the quantities given to the
hierarchy. `full_flux` retains absolute flux, whereas `spectral_shape` removes
one positive multiplicative amplitude and returns that amplitude separately.
Neither representation combines flux, kinematic moments, or SpectroPath
coordinates in one weighted vector. The existing kinematic and SpectroPath
workflows retain their own scientific meanings.

## Declared representations

Every declaration records its coordinate domain, invariances, retained and
removed quantities, required support, eligibility, units, validation domain,
and provenance. A shape representation also records a positive-homogeneous
amplitude functional,

\[
X(\xi)=\frac{F(\xi)}{A[F]}, \qquad A[aF]=aA[F],\quad a>0.
\]

`full_flux_representation()` constructs the unnormalized flux declaration.
`spectral_shape_representation()` accepts a user-supplied amplitude functional
and does not certify it. A MUSE spectrum, an ALMA frequency or velocity cube, a
NIRCam multiband vector, and another hyperspectral observation therefore need
their own amplitude definition and validation domain.

## Validated MaNGA/Sandra optical profile

`manga_sandra_spectral_shape()` fixes the configuration validated in
Experiments VIII--X. Its amplitude is the arithmetic mean of all native samples
in the complete rest-frame, vacuum-wavelength union 5050--5100, 5400--5500,
6000--6100, and 6800--6900 Angstrom. It requires a positive amplitude and
measured amplitude S/N of at least 30. Equal native samples receive equal
weight; the implementation does not replace this definition with an integral
or a resampled mean.

```{r declaration}
library(capivara)

shape <- manga_sandra_spectral_shape()
shape$validation_domain$status
shape$required_support$amplitude_windows
```

The first-order Satterthwaite statistic is available through
`spectral_shape_diagnostic()`. It diagnoses pairs under the validated
independent-Gaussian, diagonal spectral-error model. It never changes a Ward
cost and does not represent spatial covariance between MaNGA spaxels.

## Validity, support, and segmentation

Voxel validity remains a row-column-channel array. For an explicit semantic
representation, the default analysis domain is its declared measurement
validity and eligibility domain. CAPIVARA does not intersect that domain with
Starlet. Detection, host association, Starlet evidence, representation
eligibility, and final analysis support remain separate objects.

```{r semantic-workflow, eval=FALSE}
quality <- build_quality_support(
  flux = cube$imDat,
  wavelength = cube$wavelength,
  min_valid_fraction = 0.7,
  variance = variance_cube$imDat,
  dq = dq_cube,
  wavelength_interval = c(4800, 7400),
  wavelength_frame = "rest",
  interval_wavelength_frame = "rest"
)

seg <- segment_large(
  cube,
  Ncomp = 24,
  var_cube = variance_cube,
  sample_validity = quality$sample_validity,
  representation = manga_sandra_spectral_shape()
)

seg$amplitude
seg$coverage_qc
seg$hierarchy$qc
```

A declared spatial restriction remains available when the science question
requires one. Construct it explicitly with `build_capivara_support()` and pass
it as `support`; CAPIVARA then intersects it with the frozen representation
eligibility. Starlet may supply that explicit object or remain an auxiliary
multiscale diagnostic, but it is never inserted silently.

Missing channels remain absent from the transformed spectra. A merge is
admitted only when both clusters meet the reliable-overlap and feature-support
rules. Its cost is then the increase in observed-entry within-cluster SSE over
jointly represented channels. Spatial adjacency defines candidate contact but
cannot make an unsupported spectral comparison admissible.

The hierarchy preserves raw merge costs and reports inversions. Overlapping
binary64 forward-error intervals invoke the Experiment-IX secondary order: the
lexicographically smallest pair of minimum original spaxel identifiers. This
order makes unresolved numerical choices reproducible; it supplies no physical
evidence for a fixed-`Ncomp` subdivision.

## Historical calls

Calls without `representation` preserve the published median-scaled exact or
sparse workflow. That historical transformation is distinct from explicit
`full_flux`. New semantic arguments were appended to the public signatures, so
existing positional calls keep their earlier interpretation. Supplying a new
support object without a representation emits a warning because the historical
backend retains its earlier treatment of missing samples.
