---
title: "Support Masks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Support Masks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
```

A support mask identifies the spatial elements eligible for segmentation. It
does not assign region labels. Capivara can use an explicit missing-data
footprint, a white-light starlet support, or an adaptive multi-band support.

## Starlet support

Both segmentation backends accept the same starlet controls:

```{r starlet, eval=FALSE}
seg <- segment_large(
  input = cube,
  Ncomp = 25,
  use_starlet_mask = TRUE,
  support_method = "starlet",
  starlet_J = 5,
  starlet_scales = 2:5,
  include_coarse = FALSE,
  denoise_k = 0,
  starlet_mode = "soft",
  positive_only = TRUE,
  mask_mode = "na",
  knn_k = 100
)
```

The cube is collapsed to a white-light image, decomposed across starlet scales,
and reconstructed from the selected scales. `include_coarse` controls the
coarse plane; `denoise_k` is expressed in MAD units; and `positive_only`
excludes negative reconstructed values. Record every setting because a support
change alters which spaxels can receive a label.

Inspect the support independently when the footprint is scientifically
important:

```{r inspect-starlet, eval=FALSE}
support_result <- build_starlet_mask(
  cube,
  starlet_J = 5,
  starlet_scales = 2:5,
  include_coarse = FALSE,
  denoise_k = 0,
  mode = "soft",
  positive_only = TRUE
)

support_result$mask
```

## Adaptive support

The adaptive builder can combine per-band evidence and optional transforms.

```{r adaptive, eval=FALSE}
adaptive <- build_adaptive_support(
  cube,
  transform = "asinh",
  sky_method = "border",
  border_fraction = 0.1,
  z_threshold = 3,
  min_band_persistence = 2
)

seg <- segment(
  cube,
  Ncomp = 15,
  use_starlet_mask = TRUE,
  support_method = "adaptive",
  support_args = list(
    transform = "asinh",
    sky_method = "border",
    z_threshold = 3,
    min_band_persistence = 2
  )
)
```

Transform choice, sky estimator, persistence requirement, and thresholds must
be retained with the segmentation provenance. Do not describe an excluded
spaxel as a separate physical region.

## Missing and background values

With `mask_mode = "na"`, excluded cube values and returned map locations are
missing. A FITS export may convert those map values to zero for display, while
positive integers retain the Capivara region identifiers. This conversion is a
file convention; it must not be applied to the science cube silently.
